Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-16 article-models-ttinternal.tex
shortened the long proof

\begin{proof}
We first need to verify that $\jcomp{}{\epsilon_0}{\epsilon_1}$ is a morphism
from $\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}$
to $\jcomp{}{\epsilon_0}{P}$. This follows from the judgmental equality
$\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}\jdeq
\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}$, which we have proved in
\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
of exactly the right sort, so the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxext
  { \jcomp{}{\epsilon_0}{P}
    }
  { { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{P}
          }
        }
      }
    } 
  \ar{r}
    { \jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{\epsilon_1}}}}
  \ar{d}[swap]{
    \jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \idtm{
        \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
          \jcomp{}{\epsilon_1}{%
            \jcomp{}{\epsilon_0}{P}
            }
          }
        }
      }
    }
& \ctxext
    {\jcomp{}{\epsilon_0}{P}}
    {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \\
\ctxext
  {\jcomp{}{\epsilon_0}{P}}
  {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{r}[swap]{\jcomp{}{\epsilon_0}{\epsilon_1}} 
& \jcomp{}{\epsilon_0}{P}
\end{tikzcd}
\end{equation*}
commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
for the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$. Note
that this follows from \autoref{eq:extalg-eq2} provided that we can show that
\begin{align}
\jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
  \jcomp{}{\epsilon_1}{%
    \jcomp{}{\epsilon_0}{\epsilon_1}}}
& \jdeq
  \jcomp{}{\epsilon_0}{%
    \jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
    }
  \label{eq:extalg-infty1}
  \\
\jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
  \idtm{
    \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
& \jdeq
\jcomp{}{\epsilon_0}{%
  \jvcomp{}{\epsilon_1}{%
    \idtm{
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
  \label{eq:extalg-infty2}
\end{align}
Note that \autoref{eq:extalg-infty1} follows if we can show that
\begin{equation*}
\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \jdeq
  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}}.
\end{equation*}
This is a special case of \autoref{lem:extalg-twins}. The second judgmental
equality, \autoref{eq:extalg-infty2}, is trivial.
\end{proof}

>>>>

\begin{proof}
We first need to verify that $\jcomp{}{\epsilon_0}{\epsilon_1}$ is a morphism
from $\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}$
to $\jcomp{}{\epsilon_0}{P}$. This follows from the judgmental equality
$\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}\jdeq
\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}$, which we have proved in
\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
of exactly the right sort, so the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxext
  { \jcomp{}{\epsilon_0}{P}
    }
  { { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{P}
          }
        }
      }
    } 
  \ar{r}
    { \jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{\epsilon_1}}}}
  \ar{d}[swap]{
    \jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \idtm{
        \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
          \jcomp{}{\epsilon_1}{%
            \jcomp{}{\epsilon_0}{P}
            }
          }
        }
      }
    }
& \ctxext
    {\jcomp{}{\epsilon_0}{P}}
    {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \\
\ctxext
  {\jcomp{}{\epsilon_0}{P}}
  {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{r}[swap]{\jcomp{}{\epsilon_0}{\epsilon_1}} 
& \jcomp{}{\epsilon_0}{P}
\end{tikzcd}
\end{equation*}
\end{small}
commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
for the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$. Note
that this follows from \autoref{eq:extalg-eq2} provided that we can show that
\begin{align}
\jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
  \jcomp{}{\epsilon_1}{%
    \jcomp{}{\epsilon_0}{\epsilon_1}}}
& \jdeq
  \jcomp{}{\epsilon_0}{%
    \jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
    }
  \label{eq:extalg-infty1}
  \\
\jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
  \idtm{
    \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
& \jdeq
\jcomp{}{\epsilon_0}{%
  \jvcomp{}{\epsilon_1}{%
    \idtm{
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
  \label{eq:extalg-infty2}
\end{align}
Note that \autoref{eq:extalg-infty1} follows if we can show that
\begin{equation*}
\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \jdeq
  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}}.
\end{equation*}
This is a special case of \autoref{lem:extalg-twins}. The second judgmental
equality, \autoref{eq:extalg-infty2}, is trivial.
\end{proof}
